---
title: "Vaccines and Infectious Disease"
author: Aris Paschalidis
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vaccines and Infectious Disease}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(CAvaccines)
```

# Read in Data
The first step in any analysis we want to perform is to first read in our datasets.
For this analysis, we will utilize two datasets both found in the Data/ folder
of this project: pertussis and vaccination. Both of these datasets are documented
in the associated help files, but we summarize here for convenience. Pertussis 
is a dataset containing all pertussis cases, arranged by county, from the years of 
2008-2009 in the state of California. Vaccination is a dataset that contains detailed
information about vaccination rates for several thousand schools in CA.

One approach to examining our data is to study the variable names.

```{r}
# We print the first couple lines from each dataset
names(vaccination)
names(pertussis)
```

Another, likely more useful option is to visualize our data. To do this, we 
will create bar plots and density plots of our variables.

```{r}
# bar_plot(vaccination, "Jurisdiction", limits = c(0, 1600))

bar_plot(vaccination, "School_Type", limits = c(0, 6000))

density_plot(vaccination, "Enrollment")
```

