---
title: "Vaccines and Infectious Disease"
author: Aris Paschalidis
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vaccines and Infectious Disease}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(CAvaccines)
library(ggplot2)
```

# The Data
The first step in any analysis we want to perform is to first read in our 
datasets. For this analysis, we will utilize two datasets, both found in the 
Data/ folder of this project: county_vaccination, and pertussis. Both these 
datasets are documented in the associated help files, but we summarize here for
convenience. Before continuing with this vignette, we strongly encourage readers
to examine the documentation for each dataset, if they have not already.

* `county_vaccination` is a dataset containing information about vaccination rates
for each CA county for the years of 2008-2017.
* `pertussis` is a dataset containing all pertussis cases, arranged by county, from
the years of 2008-2017 in the state of California.

We first print out the variable names so that readers are aware of the variables
in each dataset.

```{r}
# We print the names of each dataset
names(county_vaccination)
names(pertussis)
```

## Plotting The Data

One particularly useful way to understand our data is to visualize it. To do 
this, we will create various plots of our variables. 

#### Incidence
We first start with our `pertussis` dataset. One way to visualize our data is
to create boxplots of the cases and rates every year.

```{r}
# Remove missing data
pertussis_omit <- na.omit(pertussis)

# Cases
ggplot(pertussis_omit, aes(x = Year, y = Cases, fill = as.numeric(Year))) + 
  geom_boxplot(outlier.shape = NA) + geom_jitter(size = 0.5, color = "gray40") +
  viridis::scale_fill_viridis() +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5, size = 15),
        axis.title = element_text(size = 10)) +
  labs(title = "Boxplot of Cases By Year", x = "Year", y = "Cases")

# Rates
ggplot(pertussis_omit, aes(x = Year, y = Rates, fill = as.numeric(Year))) + 
  geom_boxplot(outlier.shape = NA) + geom_jitter(size = 0.4, color = "gray40") +
  viridis::scale_fill_viridis() +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5, size = 15),
        axis.title = element_text(size = 10)) +
  labs(title = "Boxplot of Rates By Year", x = "Year", y = "Rates")

# Code below is a method to remove outliers if deemed needed
# ylim_cases = grDevices::boxplot.stats(pertussis_omit$Rates)$stats[c(1, 5)]
# ylim_cases = quantile(pertussis_omit$Rates, c(0, 1))
# include_cases + coord_cartesian(ylim = ylim_cases)
```

As is clear by the above graphs, in some counties the rate and cases are very
high and thus the data is hard to interpret. In order to combat this, instead
of plotting the cases and rates for every county, we could simply find the 
average number each year. This approach is shown below, 

```{r}
# Create dataframe
pertussis_means <- pertussis_omit %>% 
  dplyr::group_by(Year) %>% 
  dplyr::summarize(
    avg_cases = mean(Cases), 
    avg_rates = mean(Rates))

# Plot average for cases and rates
ggplot(pertussis_means, aes(x = Year, y = avg_cases, fill = avg_cases)) + 
  geom_col() + viridis::scale_fill_viridis() +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5, size = 15),
        axis.title = element_text(size = 10)) +
  labs(title = "Histogram of Average Cases By Year", 
       x = "Year", 
       y = "Average Cases")

ggplot(pertussis_means, aes(x = Year, y = avg_rates, fill = avg_rates)) +
  geom_col() + viridis::scale_fill_viridis() + 
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5, size = 15),
        axis.title = element_text(size = 10)) +
  labs(title = "Histogram of Average Rates By Year", 
       x = "Year", 
       y = "Average Rates")
```

To further understand what our data looks like for any particular year, we will
construct a map of the rates of infection for a typical year: 2013. Since the
data from 2013 spans a large range, we will transform the shading of the map by
log base 2. This will make it easier to interpret. Note that when we transform
the data, we do not actually change the values of our data. The data remains the
same, but we change the way data points are colored.

```{r}
# Select year 2013
pertussiss_to_map <- pertussis %>% 
  dplyr::filter(Year == "2013") 

# Map with all data from 2013
map_plot(pertussiss_to_map,
         "Rates",
         breaks = c(0, 2, 4, 8, 16, 32, 64, 128),
         title = "Rates Per 100,000 Individuals In 2013",
         fill = "Rates",
         trans = scales::pseudo_log_trans(base = 2))
```

We below present a way to create a gif from the data for all years. This is a 
superior approach to only plotting one year as was done in the previous code
chunk. Note that in order to run the below code chunk, some packages may first 
need to be installed.

```{r}
# Create dataset
pertussiss_animated <- pertussis %>% na.omit()
pertussiss_animated$Year <- as.numeric(as.character(pertussiss_animated$Year))

# Animate data
if (!requireNamespace("gganimate", quietly = TRUE)) {
  warning('Package \"gganimate\" needed in order to animate the figures. 
          Please install it. Note that user may also need to install the 
          following: "transformr", "gifski", "png".',
    call. = FALSE)
} else {
  # Animate data
  animated <- map_plot(pertussiss_animated,
                      "Rates",
                      breaks = c(0, 2, 4, 8, 16, 32, 64, 128),
                      title = "Rates Per 100,000 Individuals",
                      fill = "Rates",
                      trans = scales::pseudo_log_trans(base = 2)) +
    gganimate::transition_time(as.integer(Year)) +
    labs(title = "Rates Per 100,000 Individuals in {frame_time}")
  
  gganimate::animate(animated, fps = 5)
  # gganimate::anim_save("Animated Rates By Year.gif")
}
```

#### Vaccination Rates

We now move on to our `county_vaccination` dataset. We first start by creating 
maps of CA counties and shading the counties according to a specific 
measurement. The following set of graphs shows the percent of students with DTP 
vaccinations, the percent of students that are up to date on their vaccinations,
the percent of students that have a PME, and the percent of students that have 
a PBE.

One option for visualizing these variables is to look at a specific year. Below,
we look at the year 2013.

```{r}
# Graphs are for the year 2013
vac_to_map <- dplyr::filter(county_vaccination, Year == "2013") %>% 
  dplyr::select(Jurisdiction, percent_DTP, percent_Up_to_Date, percent_PME, percent_PBE) %>%
  dplyr::mutate_at(dplyr::vars(dplyr::contains("percent_")), ~.*100)

map_plot(vac_to_map, "percent_Up_to_Date", title = "Up to Date")
map_plot(vac_to_map, "percent_PME", title = "PME")
map_plot(vac_to_map, "percent_PBE", title = "PBE")
map_plot(vac_to_map, "percent_DTP", title = "DTP Vaccinations")
```

As we did with our `pertussis` dataset, however, we can also animate out maps.

```{r}
# Create dataset
vac_animated <- county_vaccination %>%
  dplyr::select(Jurisdiction, Year, 
                percent_DTP, percent_Up_to_Date, percent_PME, percent_PBE) %>%
  dplyr::mutate_at(dplyr::vars(dplyr::contains("percent_")), ~.*100)
vac_animated$Year <- as.numeric(as.character(vac_animated$Year))

str(vac_animated)

# Animate data
if (!requireNamespace("gganimate", quietly = TRUE)) {
  warning('Package \"gganimate\" needed in order to animate the figures. 
          Please install it. Note that user may also need to install the 
          following: "transformr", "gifski", "png".',
    call. = FALSE)
} else {
  # Animate up to data data
  animated <- map_plot(vac_animated,
                      "percent_Up_to_Date") +
    gganimate::transition_time(as.integer(Year)) +
    labs(title = "Up to Date in {frame_time}")
  
  gganimate::animate(animated, fps = 5)
  # gganimate::anim_save("Animated percent_Up_to_Date")
  
  # Animate PME data
  animated <- map_plot(vac_animated,
                      "percent_PME") +
    gganimate::transition_time(as.integer(Year)) +
    labs(title = "PME in {frame_time}")
  
  gganimate::animate(animated, fps = 5)
  # gganimate::anim_save("Animated percent_PME.gif")
  
  # Animate PBE data
  animated <- map_plot(vac_animated,
                      "percent_PBE") +
    gganimate::transition_time(as.integer(Year)) +
    labs(title = "PBE in {frame_time}")
  
  gganimate::animate(animated, fps = 5)
  # gganimate::anim_save("Animated percent_PBE.gif")
  
  # Animate DTP data
  animated <- map_plot(vac_animated,
                      "percent_DTP") +
    gganimate::transition_time(as.integer(Year)) +
    labs(title = "DTP in {frame_time}")
  
  gganimate::animate(animated, fps = 5)
  # gganimate::anim_save("Animated percent_DTP.gif")
}
```


# Modeling
We first set up data for modeling.
```{r}
`%notin%` <- Negate(`%in%`)

model_data  <- dplyr::full_join(county_vaccination, pertussis, by = c("Jurisdiction", "Year")) %>% 
  dplyr::filter(Year %notin% c("2010")) %>%
  dplyr::filter(Jurisdiction != "totals")
```

By total numbers and cases.
```{r}
scaled_total_data <- model_data %>% 
  dplyr::select(-one_of(c("Jurisdiction", "Year", "Rates"))) %>%
  dplyr::select(-contains("percent_")) %>%
  dplyr::mutate_if(is.numeric, scale) %>% 
  as.data.frame() %>% na.omit()

summary(scaled_total_data)

results_cases <- list()
for (i in 1:1000){
  t_sample = caTools::sample.split(scaled_total_data$Cases, SplitRatio = 0.7)
  t_train = subset(scaled_total_data, t_sample == TRUE)
  t_test  = subset(scaled_total_data, t_sample == FALSE)
  
  lm <- stats::lm(Cases ~., data = t_train)
  # summary(lm)
  
  predict_lm <- lm %>% stats::predict(t_test)
  results_cases <- c(results_cases, 
                     round(caret::R2(predict_lm, t_test$Cases), 4))
}
print(paste("Average Test R2 Linear:", mean(unlist(results_cases))))
```

By percents and rates
```{r}
scaled_percent_data <- model_data %>% 
  dplyr::select(-one_of(c("Jurisdiction", "Year", "Cases"))) %>%
  dplyr::select(-contains("total_")) %>%
  # dplyr::mutate_if(is.numeric, scale) %>%
  as.data.frame() %>% na.omit()

# scaled_percent_data$Rates <- scaled_percent_data$Rates/1000
summary(scaled_percent_data)

results_percent <- list()
for (i in 1:1000){
  p_sample = caTools::sample.split(scaled_percent_data$Rates, SplitRatio = 0.7)
  p_train = subset(scaled_percent_data, p_sample == TRUE)
  p_test  = subset(scaled_percent_data, p_sample == FALSE)
  
  lm <- stats::lm(Rates ~., data = p_train)
  # summary(lm)
  
  predict_lm <- lm %>% stats::predict(p_test)
  results_percent <- c(results_percent, 
                       round(caret::R2(predict_lm, p_test$Rates), 4))
}
print(paste("Average Test R2 Linear:", mean(unlist(results_percent))))
```


```{r}
corr_data <- model_data %>% 
  dplyr::select(-"Jurisdiction")
summary(corr_data)


corr_mat <- Hmisc::rcorr(as.matrix(scaled_total_data), type = "spearman")
M <- corr_mat$r
P_mat <- corr_mat$P

corrplot::corrplot(M, method = "color", col = RColorBrewer::brewer.pal(n=8, name="PuBu"),
  addCoef.col = "black",
  tl.col = "black", tl.srt = 50,tl.cex = 0.60,
  p.mat = P_mat, sig.level = 0.05, pch.cex = 5,
  mar=c(0,0,2,0), number.cex = 1)
```



