---
title: "Vaccines and Infectious Disease"
author: Aris Paschalidis
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vaccines and Infectious Disease}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(CAvaccines)
library(ggplot2)
```

# Read in Data
The first step in any analysis we want to perform is to first read in our datasets.
For this analysis, we will utilize three datasets both found in the Data/ folder
of this project: school_vaccination, county_vaccination, and pertussis. All three 
of these datasets are documented in the associated help files, but we summarize here
for convenience. Before continuing with this vignette, we strongly encourage readers
to examine the documentation for each dataset, if they have not already.

* `school_vaccination` is a dataset that contains detailed information 
about vaccination rates for several thousand schools in CA. 
* `county_vaccination` is a related dataset to school_vaccination. It
contains information about vaccination rates for each CA county.
* `pertussis` is a dataset containing all pertussis cases, arranged by 
county, from the years of 2008-2009 in the state of California.

One approach to examining our data is to study the variable names. 

```{r}
# We print the names of each dataset
names(school_vaccination)
names(county_vaccination)
names(pertussis)

# vaccine_pertussis = merge(county_vaccination, pertussis, by = "Jurisdiction")
```

Another, likely more useful option, is to visualize our data. To do this, we 
will create bar plots and density plots of our variables.

```{r}
bar_plot(school_vaccination, "Jurisdiction", limits = c(0, 1600)) + theme(legend.position = "none")

# bar_plot(school_vaccination, "School_Type", limits = c(0, 6000))
# 
# density_plot(school_vaccination, "Enrollment")
# 
# ggplot(school_vaccination, aes(x = Jurisdiction, y = PME, fill = Jurisdiction)) + geom_boxplot(color = "white") +
#     theme_dark() +
#     theme(panel.background = element_rect(fill = "#2D2D2D"),
#           plot.title = element_text(hjust = 0.5, size = 10), 
#           axis.title = element_text(size = 7),
#           axis.text.x = element_text(angle = 90),
#           legend.position = "none") +
#     scale_x_discrete(labels = pertussis$Jurisdiction) +
#     scale_y_continuous(limits = c(0, 100))
# 
# boxplot(PME ~ Jurisdiction, data=school_vaccination, xlim=c(1,59), ylim=c(0,100), xaxt='n',
#         ylab="PME % from county population", 
#         title="Permanent Medical Exemption, aggregated by counties")
# axis(side = 2)
# text(2:59, y=par("usr")[3]-0.5, labels = pertussis$Jurisdiction, srt = 90, cex = 0.6, pos = 2, xpd = TRUE)

```

